---
title: 'Statistical analysis'
author: 'Sasha D. Hafner'
date: "`r format(Sys.time(), '%H:%M - %d %B, %Y')`"
---

# Data table
```{r}
setDT(isumm)
```

# Trials 1-4 (digestate treatment effects)

Subset.

```{r}
isumm1 <- isumm[new.ID %in% as.character(1:4)]
```

Separate digestate type from treatment

```{r}
isumm1[, dig.ID := substr(treat1, 1, 1)]
isumm1[, dig.treat := gsub('^[ABC]', '', treat1)]
isumm1[dig.treat == '', dig.treat := 'none']
isumm1[, dig.treat := relevel(factor(dig.treat), ref = 'none')]
```

Fit mixed-effects model and fixed-effects for comparison.

```{r}
m0 <- lm(e.rel.150 ~ dig.treat * dig.ID + new.ID, data = isumm1)
m1 <- lmer(e.rel.150 ~ dig.treat + (1|dig.ID) + (1|new.ID), data = isumm1)
m1notreat <- lmer(e.rel.150 ~ (1|dig.ID) + (1|new.ID), data = isumm1)
m1nodig <- lmer(e.rel.150 ~ dig.treat + (1|new.ID), data = isumm1)
m2 <- lmer(e.rel.150 ~ dig.treat + (1|dig.ID/dig.treat) + (1|new.ID), data = isumm1)
m2b <- lmer(e.rel.150 ~ (1|dig.ID/dig.treat) + (1|new.ID), data = isumm1)
m3 <- lmer(e.rel.150 ~ treat1 + (1|new.ID), data = isumm1)
```

```{r}
summary(m0)
anova(m0)
```

```{r}
summary(m1)
summary(m2)
summary(m2b)
summary(m3)
AIC(m1, m2, m2b, m3)
```

Check for treatment effect and interaction with likelihood ratio test.

```{r}
anova(m1, m1notreat, test = 'Chisq')
anova(m1, m1nodig, test = 'Chisq')
anova(m2, m1, test = 'Chisq')
anova(m2b, m1, test = 'Chisq')
anova(m2b, m2, test = 'Chisq')
```

Plot to check treatment effect and interaction.

```{r}
ggplot(isumm1, aes(dig.treat, e.rel.150, colour = dig.ID)) +
  geom_jitter(height = 0) +
  theme_bw()
```

```{r}
table(isumm1[, .(dig.ID, dig.treat)])
```

Major imbalance.
Get marginal means.

```{r}
emmeans(m3, 'treat1')
mmeans1_4 <- emmeans(m3, 'treat1')
mmeans1_4
```

```{r}
pairs(mmeans1_4)
```

Get letters.

```{r}
letters1_4 <- cld(object = mmeans1_4,
                  adjust = "Tukey",
                  Letters = letters,
                  alpha = 0.05)
letters1_4
```

Sort letters by treat1 and replace mmeans.

```{r}
mmeans1_4 <- letters1_4[order(letters1_4$treat1), ]
```

And get a column for copy/paste into paper.

```{r}
mmeans1_4$tabval <- paste(round(100 * mmeans1_4$emmean, 1), gsub(' ', '', mmeans1_4$.group))
```

Residuals.

```{r}
plot(fitted(m3), resid(m3))
abline(0,0)
qqnorm(resid(m3))
qqline(resid(m3))
```

# More on trials 1-4, DM and pH effects
Look for DM and pH explanation of digestate variability.

```{r}
dref <- isumm1[dig.treat == 'none', ]
m4 <- lmer(e.rel.150 ~ (1|dig.ID) + (1|new.ID), data = dref)
m4b <- lmer(e.rel.150 ~ (1|dig.ID/new.ID), data = dref)
m5 <- lmer(e.rel.150 ~ (1|dig.ID) + (1|new.ID) + man.dm + man.ph, data = dref)
m6 <- lmer(e.rel.150 ~ (1|dig.ID/new.ID) + man.dm + man.ph, data = dref)
```

```{r}
summary(m4)
summary(m4b)
summary(m5)
summary(m6)
AIC(m4, m5)
anova(m4, m5, test = 'Chisq')
```

I do not like how AIC difference flips with ML and REML!
But see reduction in dig.ID standard deviation when DM and pH are included.
And see likelihood ratio tests below.

```{r}
m5noph <- update(m5, ~ . - man.ph)
m5nodm <- update(m5, ~ . - man.dm)
```

```{r}
m6noph <- update(m6, ~ . - man.ph)
m6nodm <- update(m6, ~ . - man.dm)
```


```{r}
summary(m5noph)
summary(m5nodm)
```

```{r}
anova(m5, m5noph, test = 'Chisq')
anova(m5, m5nodm, test = 'Chisq')
```

```{r}
anova(m6, m6noph, test = 'Chisq')
anova(m6, m6nodm, test = 'Chisq')
```

Plots to look at apparent effects.

```{r}
ggplot(dref, aes(man.dm, e.rel.150, colour = dig.ID)) +
  geom_text(aes(label = paste0(dig.ID, '-', new.ID))) +
  theme_bw()
```

```{r}
ggplot(dref, aes(man.dm, e.rel.150, colour = man.ph)) +
  geom_text(aes(label = paste0(dig.ID, '-', new.ID))) +
  theme_bw()
```

```{r}
ggplot(dref, aes(man.ph, e.rel.150, colour = dig.ID)) +
  geom_text(aes(label = paste0(dig.ID, '-', new.ID))) +
  theme_bw()
```

Take means--I do not believe pH result.

```{r}
dmn <- dref[, .(e.rel.150 = mean(e.rel.150), man.dm = mean(man.dm), man.ph = mean(man.ph)),
     by = .(dig.ID, new.ID)]
```

```{r}
m10 <- lmer(e.rel.150 ~ (1|dig.ID), data = dmn)
m10b <- lm(e.rel.150 ~ 1, data = dmn)
m11 <- lm(e.rel.150 ~ man.dm + man.ph, data = dmn)
```

```{r}
summary(m10)
summary(m10b)
summary(m11)
```

```{r}
ggplot(dmn, aes(man.dm, e.rel.150, colour = man.ph)) +
  geom_text(aes(label = paste0(dig.ID, '-', new.ID))) +
  theme_bw()
```

# Trials 5-10 (app tech on winter wheat)

Subset.

```{r}
isumm1 <- droplevels(isumm[new.ID %in% c('5', '6', '7', '8', '9', '10') & !treat1 %in% c('TH-4', 'TS1-4')])
```

```{r}
table(isumm1[, .(new.ID, treat1)])
```

Close to balanced.

```{r}
ggplot(isumm1, aes(new.ID, e.rel.150, colour = treat1)) +
  geom_jitter(height = 0) +
  theme_bw()
```

Fit mixed-effects model.

```{r}
m1 <- lmer(e.rel.150 ~ treat1 + (1|new.ID), data = isumm1)
```

```{r}
summary(m1)
```

Marginal means.

```{r}
mmeans <- emmeans(m1, 'treat1')
mmeans
```

Need name for export.

Tukey's test

```{r}
mmeans5_10 <- emmeans(m1, 'treat1')
pairs(mmeans5_10)
```

```{r}
letters5_10 <- cld(object = mmeans5_10,
                  adjust = "Tukey",
                  Letters = letters,
                  alpha = 0.05)
letters5_10
```

```{r}
mmeans5_10 <- letters5_10[order(letters5_10$treat1), ]
```

And get a column for copy/paste into paper.

```{r}
mmeans5_10$tabval <- paste(round(100 * mmeans5_10$emmean, 1), gsub(' ', '', mmeans5_10$.group))
```


# Add 11-12

```{r}
isumm1 <- droplevels(isumm[new.ID %in% c('11', '12') & treat1 != 'TS1 + acid', ])
```

```{r}
table(isumm1[, .(new.ID, treat1)])
```

Completely balanced and only two experiments, so no need for mixed-effects model.

```{r}
m1 <- lm(e.rel.150 ~ treat1 + new.ID, data = isumm1)
```

```{r}
summary(m1)
```

Marginal means.

```{r}
mmeans <- emmeans(m1, 'treat1')
mmeans
```

Need name for export.

Tukey's test

```{r}
mmeans11_12 <- emmeans(m1, 'treat1')
pairs(mmeans11_12)
```

```{r}
letters11_12 <- cld(object = mmeans11_12,
                  adjust = "Tukey",
                  Letters = letters,
                  alpha = 0.05)
```

Order.

```{r}
mmeans11_12 <- letters11_12[c(3, 2, 1), ]
```

And get a column for copy/paste into paper.

```{r}
mmeans11_12$tabval <- paste(round(100 * mmeans11_12$emmean, 1), gsub(' ', '', mmeans11_12$.group))
```

# Trials 9 and 10 for driving speed

Subset.

```{r}
isumm1 <- isumm[new.ID %in% as.character(c(9, 10))]
```

Extract driving speed and application method.

```{r}
table(isumm1$treat)
isumm1[, speed := factor('12', levels = c(4, 12))]
isumm1[grepl('-4', treat1), speed := '4']
```

```{r}
isumm1[, app.mthd := factor(sub('-4', '', treat1), levels = c('TH', 'TS1'))]
```

```{r}
table(isumm1[, .(new.ID, app.mthd, speed)]) 
```

Completely crossed, perfectly balanced.

Plot.

```{r}
ggplot(isumm1, aes(as.numeric(speed), e.rel.150, colour = interaction(app.mthd, new.ID), shape = new.ID, lty = new.ID)) +
	geom_smooth(method = lm, se = FALSE) +
	geom_jitter(height = 0)
```


Fit model.

```{r}
m1 <- lm(log10(e.rel.150) ~ app.mthd * speed * new.ID, data = isumm1)
m2 <- aov(log10(e.rel.150) ~ (app.mthd + speed + new.ID)^2, data = isumm1)
m3 <- aov(e.rel.150 ~ (app.mthd + speed + new.ID)^2, data = isumm1)
```

```{r}
summary(m1)
anova(m1)
drop1(m1, test = 'F')
drop1(m2, test = 'F')
drop1(m3, test = 'F')
summary(m3, test = 'F')
```

```{r}
plot(m1, ask = FALSE)
```

```{r}
model.tables(m2, 'means')
```

More useful means.

```{r}
isumm1[, .(mn = mean(e.rel.150), sd = sd(e.rel.150)), by = .(new.ID, app.mthd, speed)]
```

# Trials 13 and 14 for hose distance effect

Subset.

```{r}
isumm1 <- isumm[new.ID %in% as.character(c(13, 14))]
```

Extract separation status and hose distance.

```{r}
table(isumm1$treat)
class(isumm1)
isumm1[, treat1 := tolower(treat1)]
isumm1[, distancef := factor(sub('sep|un', '', treat1))]
isumm1[, distancen := as.numeric(sub('sep|un', '', treat1))]
isumm1[, sep := factor(sub('[0-9]+$', '', treat1))]
```

```{r}
table(isumm1[, .(new.ID, sep, distancef)]) 
```

Subset is completely crossed, perfectly balanced.
But `un` has an additional distance of 12 cm.

Plot.

```{r}
ggplot(isumm1, aes(as.numeric(distancen), e.rel.150, colour = interaction(sep, new.ID), shape = new.ID, lty = new.ID)) +
	geom_smooth(method = lm, se = FALSE) +
	geom_point(height = 0)
```

Hmm, might be some interactions.
Fit model.

```{r}
m1 <- lm(e.rel.150 ~ sep * distancen * new.ID, data = isumm1)
m2 <- lm(e.rel.150 ~ (sep + distancen + new.ID)^2, data = isumm1)
m2b <- lm(e.rel.150 ~ (sep + distancen + new.ID)^2, data = isumm1, subset = sep != 'sep' | new.ID != 13)
```

```{r}
summary(m1)
anova(m1)
drop1(m1, test = 'F')
drop1(m2, test = 'F')
summary(m2b)
```

No clear distance effects, marginal interaction evidence, but all related to those negative values.

```{r}
plot(m1, ask = FALSE)
```

Means.

```{r}
isumm1[, .(mn = mean(e.rel.150), sd = sd(e.rel.150)), by = .(new.ID, sep, distancef)]
```

# Trial 15, chamber movement

Subset.

```{r}
isumm1 <- isumm[new.ID == 15, ]
```

Set reference.

```{r}
isumm1[, treat1 := factor(treat1, levels = c('No', '2-Pos', '7-Pos'))]
```

Fit model.

```{r}
m1 <- aov(e.rel.150 ~ treat1, data = isumm1)
```

Tukey's test.

```{r}
summary(m1)
TukeyHSD(m1)
```

Dunnett test.

```{r}
mmeans15 <- emmeans(m1, 'treat1')
mmeans15
contrast(mmeans15, 'trt.vs.ctrl')
contrast(mmeans15, 'tukey')
pairs(mmeans15)
```

Relative effects.

```{r}
coef(m1)
coef(m1) / coef(m1)[1]
```

```{r}
plot(m1)
```



# Other stuff
I still need to sort this stuff out!

Then temperature stuff
Add to supporting material
