---
title: 'Linear model analysis for temperature difference size'
output: pdf_document
author: 
date: "`r Sys.time()`"
---

```{r}
library(viridisLite)
```

Temperature difference.

```{r}
d2$dtemp <- d2$Itemp - d2$Otemp
```

Check interpretation.

```{r}
d2[, plot := factor(plot)]
```

Should have 0 or 1 below.

```{r}
table(table(d2[, .(plot, t.start)]))
```

Looks good.

```{r}
fit1 <- lm(Itemp ~ poly(glorad, 3) + poly(Otemp, 3) + poly(wv2, 3) + plot, data = d2)
fit2 <- lm(dtemp ~ poly(glorad, 3) + poly(Otemp, 3) + poly(wv2, 3) + plot, data = d2)
fit2e <- lm(dtemp ~ glorad + I(glorad^2) + I(glorad^3) + Otemp + I(Otemp^2) + I(Otemp^3) + 
	    wv2 + I(wv2^2) + I(wv2^3) + plot, data = d2)
```

Check results

```{r}
summary(fit1)
summary(fit2)
summary(fit2e)
```

The two models are identical.
(R-squared is lower for dtemp but that is just because we have already removed a lot of the variation by calculating a difference.)
Interesting and reassuring that `plot` effects are small.

Let's generate scaled predictor variables for standardized coefficients (relative to 1 standard deviation of predictor variable).
This will show which predictors are the most important compared to how much they vary.

```{r}
fit3 <- lm(Itemp ~ poly(scale(glorad), 3) + poly(scale(Otemp), 3) + poly(scale(wv2), 3) + plot, data = d2)
```

```{r}
summary(fit3)
```

It looks like temperature (`Otemp`) is the most important.
Is that supported by the measurements?

```{r}
pairs(d2[, .(Otemp, glorad, wv2, dtemp)])
```

Seem so, yes.

Let's see how much worse the model is without the other two.

```{r}
fit4 <- lm(Itemp ~ Otemp + I(Otemp^2) + I(Otemp^3)  + plot, data = d2)
```

```{r}
summary(fit4)
summary(fit1)
```

Worse than 2, not that much though.
So effects of radiation and wind are not huge.

Generate predictions for plotting.

```{r}
d2$Itemp.pred <- predict(fit1)
d2$Itemp.pred2 <- predict(fit2)
d2$Itemp.pred4 <- predict(fit4)
```

And take a look.

```{r}
ggplot(d2, aes(Otemp, Itemp, colour = glorad)) + 
  geom_line(aes(Otemp, Itemp.pred), colour = 'blue') + 
  geom_line(aes(Otemp, Itemp.pred4), colour = 'red') + 
  geom_point() + 
  scale_color_viridis_c(option = 'magma') +
  theme_bw() +
  xlab('Ambient soil surface temperature (C)') + ylab('DFC soil surface temperature (C)')
```

So, temperature alone indeed does as well as the most complete model.
But both miss a lot of the variation.
How about effects of earlier weather?
For that we need to add lagged predictor variables.
Try previous hour.

```{r}
wthr <- unique(d2[, .(t.start, Otemp, glorad, wv2)])
wthr[, t.start := t.start - 3600]
d2.orig <- d2
d2 <- merge(d2, wthr, by = 't.start', suffixes = c('', '.lag1'))
wthr[, t.start := t.start - 3600]
d2 <- merge(d2, wthr, by = 't.start', suffixes = c('', '.lag2'))
```

First, try interactions.
Then include lagged predictors.

```{r}
fit5 <- lm(Itemp ~ (poly(glorad, 3) + poly(Otemp, 3) + poly(wv2, 3))^2 + plot, data = d2)
fit6 <- lm(Itemp ~ (poly(glorad, 3) + poly(Otemp, 3) + poly(wv2, 3) + 
	   poly(glorad.lag1, 3) + poly(Otemp.lag1, 3) + poly(wv2.lag1, 3) + 
	   poly(Otemp.lag2, 3))^2 + plot, data = d2)
summary(fit5)
summary(fit6)
summary(fit2)
```

A lot of improvement (with interactions included).

```{r}
d2$Itemp.pred5 <- predict(fit5)
d2$Itemp.pred6 <- predict(fit6)
```

```{r}
ggplot(d2, aes(Otemp, Itemp, colour = glorad)) + 
  geom_line(aes(Otemp, Itemp.pred), colour = 'blue') + 
  geom_line(aes(Otemp, Itemp.pred4), colour = 'red') + 
  geom_line(aes(Otemp, Itemp.pred5), colour = 'orange') + 
  geom_line(aes(Otemp, Itemp.pred6), colour = 'skyblue') + 
  geom_point() + 
  scale_color_viridis_c(option = 'magma') +
  theme_bw() +
  xlab('Ambient soil surface temperature (C)') + ylab('DFC soil surface temperature (C)')
```

```{r}
ggplot(d2) +
  geom_point(aes(Itemp, Itemp.pred), colour = 'blue', alpha = 0.2) +
  geom_point(aes(Itemp, Itemp.pred4), colour = 'red', alpha = 0.2) +
  geom_point(aes(Itemp, Itemp.pred5), colour = 'orange', alpha = 0.2) +
  theme_bw() 
```

```{r}
#newdat <- expand.grid(glorad = 0:10 * 100, Otemp = 0:10 * 3, wv2 = 2)
#setDT(newdat)
#newdat$dtemp.pred5 <- predict(fit2, newdata = newdat)
#newdat[dtemp.pred5 > 5, ]
```

Looks at some subsets.

```{r}
#newdat[dtemp.pred5 > 3, ]
#newdat[dtemp.pred5 > 5, ]
#newdat[dtemp.pred5 > 8, ]
```

How common were the high temperatures and temperature differences?

```{r}
quantile(d2.orig$Otemp)
quantile(d2.orig$Itemp)
quantile(d2.orig$dtemp)
```

Mean difference by plot.

```{r}
summ <- d2[, .(dtemp.mean = mean(dtemp), dtemp.med = median(dtemp)), by = plot]
summ
```

```{r}
t.test(summ$dtemp.med)
t.test(summ$dtemp.mean)
```

Quantiles above and below 20C.
And make sure each plot (mostly) has some high temperature measurements.

```{r}
d2.orig[, .(quantile(dtemp)), by = Otemp < 20]
d2.orig[, .N, by = Otemp < 20]
d2.orig[, .N, by = .(plot, Otemp < 20)]
822 / 6418
```

```{r}
ggplot(d2, aes(Otemp, Itemp, colour = plot)) +
  geom_abline(intercept = 0, slope = 1, colour = 'gray55', lty = 2) +
  geom_point() +
  coord_fixed() +
  theme_bw() +
  labs(x = expression('Ambient soil surface temperature'~(degree*C)),
       y = expression('DFC soil surface temperature'~(degree*C)),
       colour = 'Radiation')
```








