---
title: 'Simple models'
author: 'Sasha D. Hafner'
date: "`r format(Sys.time(), '%H:%M - %d %B, %Y')`"
---

# Structure/replication

Check for replication by slurry ID x trial ID.
And number of digestate and raw slurries.

```{r}
table(ds[, .(slurry.ID, trial.ID)])
table(ds[, .(slurry.type, trial.ID)])
```

# Exploratory plots and correlation

Scatterplots to look for correlation between emission and predictor variables and between predictor variables.

```{r}
pairs(ds[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.150)], col = ds$col, pch = 19)
```

Pearson's r.
`dsd` is digestate only and `dsr` is raw slurry only.
`ds` is everything.

```{r}
cor(ds[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.150)])
cor(dsd[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.150)])
cor(dsr[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.150)])
```

Extract values for correlation with DM

```{r}
cor(ds[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.150)])[, 'TS.mn']
```



# Models - combined slurry types

## A. Slurry type models

```{r}
lm_a1 <- lm(e.rel.150 ~ slurry.type, data = ds)
summary(lm_a1)
drop1(lm_a1, test = 'F')
```

```{r}
mm_a1 <- lmer(e.rel.150 ~ slurry.type + (1|trial.ID) + (1|slurry.ID), data = ds, REML = FALSE)
summary(mm_a1)
fixef(mm_a1)
drop1(mm_a1, test = 'Chisq')
r2(mm_a1)
```

Exclude intercept to get means.

```{r}
mm_a2 <- lmer(e.rel.150 ~ slurry.type + (1|trial.ID) + (1|slurry.ID) - 1, data = ds, REML = FALSE)
summary(mm_a2)
fixef(mm_a2)
drop1(mm_a2, test = 'Chisq')
r2(mm_a2)
```

Save to list.

```{r}
modlist <- list()
modlist[['a']] <- mm_a2
```

```{r}
lm_a2 <- lm(e.rel.150 ~ slurry.type + trial.ID, data = ds)
summary(lm_a2)
drop1(lm_a2, test = 'F')
```

```{r}
lm_a3 <- aov(e.rel.150 ~ slurry.type, data = ds)
summary(lm_a3)
model.tables(lm_a3, type = 'means')
TukeyHSD(lm_a3, 'slurry.type')
```

## B. DM and pH
Hypothesize that DM and pH can explain emission.
`mm1.mn` is the only other variable not highly correlated with DM, so we include that too.


```{r}
mm_b1 <- lmer(e.rel.150 ~ TS.mn + pH.lab.mn + mm1.mn + (1|trial.ID) + (1|slurry.ID), data = ds, REML = FALSE)
summary(mm_b1)
fixef(mm_b1)
drop1(mm_b1, test = 'Chisq')
r2(mm_b1)
```

Drop slurry ID.

```{r}
mm_b2 <- lmer(e.rel.150 ~ TS.mn + pH.lab.mn + mm1.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_b2)
fixef(mm_b2)
drop1(mm_b2, test = 'Chisq')
r2(mm_b2)
```

No evidence of `mm1.mn` correlation so we'll drop it.

```{r}
mm_b3 <- lmer(e.rel.150 ~ TS.mn + pH.lab.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_b3)
fixef(mm_b3)
drop1(mm_b3, test = 'Chisq')
r2(mm_b3)
```

Save.

```{r}
modlist[['b']] <- mm_b3
```

Pretty clear, and positive coefficients.

Check linear model.

```{r}
lm_b3 <- lm(e.rel.150 ~ TS.mn + pH.lab.mn, data = ds)
summary(lm_b3)
```

Same.

Add slurry type.

```{r}
mm_b4 <- lmer(e.rel.150 ~ TS.mn + pH.lab.mn + slurry.type + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_b4)
fixef(mm_b4)
drop1(mm_b4, test = 'Chisq')
anova(mm_b4, mm_b3, test = 'Chisq')
r2(mm_b4)
AIC(mm_b4)
AIC(mm_b3)
model_performance(mm_b4, estimator = 'ML')
model_performance(mm_b3, estimator = 'ML')
```

Save

```{r}
modlist[['b+']] <- mm_b4
```

And try linear model with slurry type too.

```{r}
lm_b4 <- lm(e.rel.150 ~ TS.mn + pH.lab.mn + slurry.type, data = ds)
summary(lm_b4)
drop1(lm_b4, test = 'F')
```

Yes it does, but the fit does not improve much.

```{r}
summary(lm_f1)[c('r.squared', 'adj.r.squared')]
summary(lm_b3)[c('r.squared', 'adj.r.squared')]
summary(lm_b4)[c('r.squared', 'adj.r.squared')]
```

Similar model 

```{r}
lm_b5 <- aov(e.rel.150 ~ TS.mn + pH.lab.mn + slurry.type, data = ds)
pairs(emmeans(lm_b5, ~ slurry.type), adjust = 'tukey')
predict(lm_b5, newdata = data.frame(TS.mn = mean(ds$TS.mn), pH.lab.mn = mean(ds$pH.lab.mn), slurry.type = c('Cattle', 'Pig', 'Digestate')))
```


# C. ESA based models

```{r}
mm_c1 <- lmer(e.rel.150 ~ area.perc.mn + pH.lab.mn + mm1.mn + mm2.mn + (1|trial.ID) + (1|slurry.ID), data = ds, REML = FALSE)
summary(mm_c1)
fixef(mm_c1)
drop1(mm_c1, test = 'Chisq')
r2(mm_c1)
```

Drop slurry ID.

```{r}
mm_c2 <- lmer(e.rel.150 ~ area.perc.mn + pH.lab.mn + mm1.mn + mm2.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_c2)
fixef(mm_c2)
drop1(mm_c2, test = 'Chisq')
r2(mm_c2)
```

Go to `lm()` with no trial ID.


```{r}
lm_c3 <- lm(e.rel.150 ~ area.perc.mn + pH.lab.mn + mm1.mn + mm2.mn, data = ds)
summary(lm_c3)
drop1(lm_c3, test = 'F')
```

Try ESA alone.

```{r}
mm_c4 <- lmer(e.rel.150 ~ area.perc.mn + pH.lab.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_c4)
fixef(mm_c4)
drop1(mm_c4, test = 'Chisq')
r2(mm_c4)
```

```{r}
modlist[['c']] <- mm_c4
```

```{r}
lm_c4 <- lm(e.rel.150 ~ area.perc.mn + pH.lab.mn, data = ds)
summary(lm_c4)
drop1(lm_c4, test = 'F')
```

We get a clear correlation and a decent fit here.
Of course the sign of the slope makes no sense.
Overall fit is identical to DM based model.

```{r}
summary(lm_a3)[c('r.squared', 'adj.r.squared')]
summary(lm_c4)[c('r.squared', 'adj.r.squared')]
```

Add slurry type

```{r}
mm_c5 <- lmer(e.rel.150 ~ area.perc.mn + pH.lab.mn + slurry.type + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_c5)
drop1(mm_c5, test = 'Chisq')
r2(mm_c5)
```

```{r}
modlist[['c+']] <- mm_c5
```

# D. logK-based models

```{r}
mm_d1 <- lmer(e.rel.150 ~ logK + pH.lab.mn + mm1.mn + (1|trial.ID) + (1|slurry.ID), data = ds, REML = FALSE)
summary(mm_d1)
fixef(mm_d1)
drop1(mm_d1, test = 'Chisq')
r2(mm_d1)
```

Drop slurry.ID.

```{r}
mm_d2 <- lmer(e.rel.150 ~ logK + pH.lab.mn + mm1.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_d2)
fixef(mm_d2)
drop1(mm_d2, test = 'Chisq')
r2(mm_d2)
```

Particle predictor mm1.mn does not do much.

```{r}
mm_d3 <- lmer(e.rel.150 ~ logK + pH.lab.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_d3)
fixef(mm_d3)
drop1(mm_d3, test = 'Chisq')
r2(mm_d3)
```

Save 

```{r}
modlist[['d']] <- mm_d3
```

And logK is not very useful.

Add slurry type

```{r}
mm_d4 <- lmer(e.rel.150 ~ logK + pH.lab.mn + slurry.type + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_d4)
fixef(mm_d4)
drop1(mm_d4, test = 'Chisq')
r2(mm_d4)
```


```{r}
modlist[['d+']] <- mm_d4
```


# E. m/n models

```{r}
mm_e1 <- lmer(e.rel.150 ~ pH.lab.mn + n.mn + mm1.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_e1)
fixef(mm_e1)
drop1(mm_e1, test = 'Chisq')
r2(mm_e1)
```

Drop mm1.mn

```{r}
mm_e2 <- lmer(e.rel.150 ~ pH.lab.mn + n.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_e2)
fixef(mm_e2)
drop1(mm_e2, test = 'Chisq')
r2(mm_e2)
```

```{r}
modlist[['e']] <- mm_e2
```

Add slurry type

```{r}
mm_e3 <- lmer(e.rel.150 ~ pH.lab.mn + n.mn + slurry.type + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_e3)
drop1(mm_e3, test = 'Chisq')
r2(mm_e3)
```

```{r}
modlist[['e+']] <- mm_e3
```

```{r}
lm_e3 <- lm(e.rel.150 ~ pH.lab.mn + n.mn, data = ds)
summary(lm_e3)
drop1(lm_e3, test = 'F')
```

This is best model so far, and simple.

Does slurry type add more info?


```{r}
lm_e4 <- lm(e.rel.150 ~ pH.lab.mn + n.mn + slurry.type, data = ds)
summary(lm_e4)
drop1(lm_e4, test = 'F')
```

Yes.


# F. Particle based models

```{r}
mm_f1 <- lmer(e.rel.150 ~ pH.lab.mn + mm1.mn + mm2.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_f1)
fixef(mm_f1)
drop1(mm_f1, test = 'Chisq')
r2(mm_f1)
```

```{r}
mm_f2 <- lmer(e.rel.150 ~ pH.lab.mn + mm2.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_f2)
fixef(mm_f2)
drop1(mm_f2, test = 'Chisq')
r2(mm_f2)
```

Save 

```{r}
modlist[['f']] <- mm_f2
```

```{r}
lm_f3 <- lm(e.rel.150 ~ pH.lab.mn + mm2.mn, data = ds)
summary(lm_f3)
drop1(lm_f3, test = 'F')
```

Another good model.


Does slurry type add more info?

```{r}
lm_f4 <- lm(e.rel.150 ~ pH.lab.mn + mm2.mn + slurry.type, data = ds)
summary(lm_f4)
drop1(lm_f4, test = 'Chisq')
```

Nope, not clearly at least!

```{r}
mm_f5 <- lmer(e.rel.150 ~ pH.lab.mn + mm2.mn + slurry.type + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_f5)
fixef(mm_f5)
drop1(mm_f5, test = 'Chisq')
```

Neat!

```{r}
modlist[['f+']] <- mm_f5
```


## Model predictions

Make predictions for plots.

```{r}
ds[, mm_b3.pred := predict(mm_b3)]
ds[, mm_a1.pred := predict(mm_a1)]
```

## Table of models

```{r}
fit1 <- as.data.table(do.call(rbind, lapply(modlist[ww <- c('a', 'b', 'd', 'e', 'f')], model_performance, estimator = 'ML')))
fit1[, model := ww]
fit1
```

```{r}
fit2 <- as.data.table(do.call(rbind, lapply(modlist[ww <- c('c')], model_performance, estimator = 'ML')))
fit2[, model := ww]
fit2
```

```{r}
fit3 <- as.data.table(do.call(rbind, lapply(modlist[ww <- c('b+', 'c+', 'd+', 'e+', 'f+')], model_performance, estimator = 'ML')))
fit3[, model := ww]
fit3
```

## Summaries for tables

Model summaries

```{r}
lapply(modlist[ww <- c('a', 'b', 'c', 'd', 'e', 'f')], summary)
lapply(modlist[ww <- c('b+', 'c+', 'd+', 'e+', 'f+')], summary)
```

LRTs

```{r}
lapply(modlist[ww <- c('a', 'b', 'c', 'd', 'e', 'f')], drop1, test = 'Chisq')
lapply(modlist[ww <- c('b+', 'c+', 'd+', 'e+', 'f+')], drop1, test = 'Chisq')
```

# Predicted values for all models

```{r]
dpred <- data.table()
for (i in modlist) {
  dd <- ds 
  dd[, e.rel.pred := predict(i)]
  dpred <- rbind(dpred, dd)
}
```
