---
title: 'Predictors for 8 hr emission'
author: 'Sasha D. Hafner'
date: "`r format(Sys.time(), '%H:%M - %d %B, %Y')`"
---

# Exploratory plots and correlation

Scatterplots to look for correlation between emission and predictor variables and between predictor variables.

```{r}
pairs(ds[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.8, e.rel.150)], col = ds$col, pch = 19)
```

How does 8 h emission compare to 150 h?

```{r}
ggplot(ds, aes(e.rel.8, e.rel.150, colour = slurry.type)) +
  geom_smooth(aes(group = 1), se = FALSE, colour = 'gray65') +
  geom_point(size = 3) +
  theme_bw() 
```

Some curvature (high 8 hour stops leading to higher and higher 150 h emission) mostly driven by digestate (or 2 pig points, depending on perspective, but intercept of 0 should be assumed).

Pearson's r.
`dsd` is digestate only and `dsr` is raw slurry only.
`ds` is everything.

```{r}
cor(ds[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.8)])
cor(dsd[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.8)])
cor(dsr[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.8)])
```

Extract values for correlation with DM

```{r}
cor(ds[, .(pH.lab.mn, TS.mn, area.perc.mn, logK, n.mn, mm1.mn, mm2.mn, e.rel.8)])[, 'TS.mn']
```

## B. DM and pH
Hypothesize that DM and pH can explain emission.
`mm1.mn` is the only other variable not highly correlated with DM, so we include that too.


```{r}
mm_b1 <- lmer(e.rel.8 ~ TS.mn + pH.lab.mn + mm1.mn + (1|trial.ID) + (1|slurry.ID), data = ds, REML = FALSE)
summary(mm_b1)
fixef(mm_b1)
drop1(mm_b1, test = 'Chisq')
r2(mm_b1)
```

Drop slurry ID.

```{r}
mm_b2 <- lmer(e.rel.8 ~ TS.mn + pH.lab.mn + mm1.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_b2)
fixef(mm_b2)
drop1(mm_b2, test = 'Chisq')
r2(mm_b2)
```

No evidence of `mm1.mn` correlation so we'll drop it.

```{r}
mm_b3 <- lmer(e.rel.8 ~ TS.mn + pH.lab.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_b3)
fixef(mm_b3)
drop1(mm_b3, test = 'Chisq')
r2(mm_b3)
```

No DM effect either!
Like reviewer 1 proposed.

```{r}
mm_b4 <- lmer(e.rel.8 ~ pH.lab.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_b4)
fixef(mm_b4)
drop1(mm_b4, test = 'Chisq')
r2(mm_b4)
```

Save.

```{r}
modlist[['b']] <- mm_b4
```

# C. ESA based models

```{r}
mm_c1 <- lmer(e.rel.8 ~ area.perc.mn + pH.lab.mn + mm1.mn + mm2.mn + (1|trial.ID) + (1|slurry.ID), data = ds, REML = FALSE)
summary(mm_c1)
fixef(mm_c1)
drop1(mm_c1, test = 'Chisq')
r2(mm_c1)
```

Drop slurry ID.

```{r}
mm_c2 <- lmer(e.rel.8 ~ area.perc.mn + pH.lab.mn + mm1.mn + mm2.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_c2)
fixef(mm_c2)
drop1(mm_c2, test = 'Chisq')
r2(mm_c2)
```

Go to `lm()` with no trial ID.


```{r}
lm_c3 <- lm(e.rel.8 ~ area.perc.mn + pH.lab.mn + mm1.mn + mm2.mn, data = ds)
summary(lm_c3)
drop1(lm_c3, test = 'F')
```

Try ESA and pH.

```{r}
mm_c4 <- lmer(e.rel.8 ~ area.perc.mn + pH.lab.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_c4)
fixef(mm_c4)
drop1(mm_c4, test = 'Chisq')
r2(mm_c4)
```

```{r}
modlist[['c']] <- mm_c4
```

# D. logK-based models

```{r}
mm_d1 <- lmer(e.rel.8 ~ logK + pH.lab.mn + mm1.mn + (1|trial.ID) + (1|slurry.ID), data = ds, REML = FALSE)
summary(mm_d1)
fixef(mm_d1)
drop1(mm_d1, test = 'Chisq')
r2(mm_d1)
```

Drop slurry.ID.

```{r}
mm_d2 <- lmer(e.rel.8 ~ logK + pH.lab.mn + mm1.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_d2)
fixef(mm_d2)
drop1(mm_d2, test = 'Chisq')
r2(mm_d2)
```

Drop mm1.mn 

```{r}
mm_d3 <- lmer(e.rel.8 ~ logK + pH.lab.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_d3)
fixef(mm_d3)
drop1(mm_d3, test = 'Chisq')
r2(mm_d3)
```

Drop logK, back to pH model!

# E. m/n models

```{r}
mm_e1 <- lmer(e.rel.8 ~ pH.lab.mn + n.mn + mm1.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_e1)
fixef(mm_e1)
drop1(mm_e1, test = 'Chisq')
r2(mm_e1)
```

Drop mm1.mn

```{r}
mm_e2 <- lmer(e.rel.8 ~ pH.lab.mn + n.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_e2)
fixef(mm_e2)
drop1(mm_e2, test = 'Chisq')
r2(mm_e2)
```

Drop n.mn gets us back to pH model!

# F. Particle based models

```{r}
mm_f1 <- lmer(e.rel.8 ~ pH.lab.mn + mm1.mn + mm2.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_f1)
fixef(mm_f1)
drop1(mm_f1, test = 'Chisq')
r2(mm_f1)
```

```{r}
mm_f2 <- lmer(e.rel.8 ~ pH.lab.mn + mm2.mn + (1|trial.ID), data = ds, REML = FALSE)
summary(mm_f2)
fixef(mm_f2)
drop1(mm_f2, test = 'Chisq')
r2(mm_f2)
```

Drop mm2.mn, back at pH model yet again

